### 2024-05-28 index.qmd
Started outlining updated intro for Dourados Reserve

---
title: Epidemiological dynamics in idealized metapopulations defined by asymmetric homophily
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability
    roles: writing, model development
    corresponding: true
    email: maturner@stanford.edu
bibliography: [this.bib, ../library.bib]
# bibliography: this.bib
classoption: abstract
keep-tex: true
abstract: |
  \noindent The COVID-19 pandemic disproportionately affected economically or otherwise marginalized groups, including Indigenous groups. In the Dourados Indigenous Reserve in Mato Grosso do Sul, Brazil, it was hypothesized that young men were largely responsible for transmission in and out of the Reserve when they traveled to the neighboring City of Dourados for economic reasons. Here we develop an agent-based compartmental model of disease transfer between two groups representing the Reserve and City populations with smaller and larger population-size groups of agents, respectively. We consider heterogeneous group-level mixing parameters, where the young Indigenous men have different mobility than other Reserve residents, both of which are different from city resident mobility between Reserve and City. We test whether greater mobility among young men could account for observed variant diffusion into the Reserve from the City. We also evaluate the potential efficacy of temporary, well-timed economic relief targeted to young men in the Reserve so they would not need to travel the City. This work then demonstrates a potential real-world intervention to be used for the next pandemic. It also contributes to the theory of epidemics in metapopulations. Such scientific predictions are essential for designing a suite of equitable, effective interventions for preventing or mitigating the next pandemic.
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
---

## Introduction

To equitably prepare for pandemics in our uncertain world we need to [@Rodo2021; @Adashi2022]. The global COVID-19 pandemic highlighted the current lack of outcome equality, where a country's inequality (measured by the Gini index) predicted higher burden of COVID-19 infections [@Su2022]. In this work we contribute to more equitable pandemic preparedness by developing an agent-based compartmental model of COVID-19 transmission between a Brazilian city, Dourados, in Mato Grosso do Sul, and the neighboring Dourados Indigenous Reserve [@DeOliveira2023]. COVID-19 variant diffusion from Dourados City into the Dourados Reserve was hypothesized to be driven by young Indigenous men traveling between the reserve and the city of Dourados for economic reasons. COVID-19 burden was felt disproportionately by Indigenous peoples in Brazil. Therefore this case study demonstrates how we can begin to develop scientific predictive models of inter-community disease transmission that could be used to optimize the timing of public health interventions to limit the spread of fitter pathogen variants.

<!-- To complement our biological understanding and preventative tools for mitigating the risk of pandemics, social science is essential to understand how social structure and behavior determines how disease and disease prevention behaviors (such as vaccination[@Amin2017; @Omer2021]) are adopted by individuals in different group-structured populations [@Buckee2021; @Epstein2008; @SmaldinoJones2021]. Oftentimes in epidemiology, groups are defined by any different group partitions can be relevant to epidemiology, including age (RSV and Polio refs), occupation (frontline/remote refs), and, of course, political party [@VanBavel2020]. To understand how social structure affects . Using this we can infer the effect of different vaccine distribution allotments between the two groups that could possibly better protect one group or the other, or meet other optimal criteria. Computational analyses of metapopulation models can be used to create in silico experiments that would be impossible to implement in the real world, for instance to predict how different vaccination distribution strategies would best prevent transmission dynamics [@Singer2022]. -->

<!-- Whether a new viral mutation or a cultural adaptation, beneficial innovations diffuse differently through different metapopulation structures. Turner and colleagues (2023) showed that novel innovations with higher fitness than an alternative best diffused when first introduced to a model minority group compared to a majority group (in terms of population size), in a metapopulation with a single other majority group [@Turner2023], and diffusion was further optimized when majority groups moderately preferred in-group interactions. Preference for in- versus out-group interactions is known as homophily [@McPherson2001], and introduced to a simple model metapopulation composed of a minority and majority group (by population size), where there exists a lower-fitness alternative, fitness-biased social learning.-->

<!-- Simple idealized models like we develop here serve as a theoretical ground truth against which we can better understand real-world complexities. For instance, schistosomaiasis hotspots can be predicted with some accuracy based on data from initial outbreak [@Singer2024], but metapopulation structure may have an underlying effect on the emergence of hotspots that, when accounted for, could increase predictive power of data-driven statistical models. Metapopulation structure with asymmetric homophily is not yet rigorously understood in an idealized theoretical setting, despite the fact that contact structure can be observed to affect transmission dynamics of sexually transmitted disease [@Morris1991], the spread of antibiotic-resistant bacteria across hospital wards [@Shapiro2020], the emergence and diffusion of drug-resistant malaria [@Ariey2003], the spread of schistosomiasis [@Ciddio2017], and provide disease reservoirs that drove the spread of tuberculosis and COVID-19 between some prison and community populations [@Sacchi2015; @Ryckman2021]. Contact structure has been rigorously shown to shape epidemiological dynamics [@Arthur2017; @Großmann2021], but not in settings where social networks are defined by asymmetric homophily, specifically. For data-driven applications, asymmetric homophily measurements may be easier or more reliable compared to the network structure itself. -->

<!-- To understand the effect of asymmetric homophily levels on epidemiological dynamics, we developed a computational model where metapopulations are composed of two groups, a minority and majority group defined by population size, where group members interact more or less frequently depending on group membership and group-level homophily levels, and whether interactions are structured by a social network (more like in a sexually-transmitted disease) or by random asymmetric homophilous mixing within and between groups (like in an airborne disease), adapting the models of Turner et al (2023) and Turner, Reynolds, and Jones (2024) for epidemiology by augmenting the social learning dynamic with SIR epidemiological dynamics. -->

<!-- Models like this enable us to rigorously understand how one important mechanism affects the course of disease and optimal intervention strategy. The cost of rigor, at this stage, is the exclusion of other important details that can affect disease or cultural transmission, including individual-level general health, access to clean running water, working conditions, and social factors such as asymmetric outgroup aversion levels between groups, which can undermine the adoption of adaptive behaviors [@EhretConstantino2022]. However, this initial mechanistic model ideally sets in motion a series of follow-on studies to better understand which of these additional factors alter model predictions to what extent. Model refinements and specializations can be added for specific epidemiological contexts, different scenarios and intervention strategies can be tested computationally for optimal policies given available information. More shortcomings, refinements, and specializations can follow that, with the process repeating as necessary. This work complements a broader effort across evolutionary sciences to understand the interplay between social structure and collective co-adaptation among humans, non-humans, and pathogens [@Bak-Coleman2021; @GalesicEtAl2023]. -->

## Model

-   SIR model with

## Analysis

## Discussion

##