### 2024-07-14 index.qmd
updated index and restarted Rproj

---
title: Epidemiological dynamics in idealized metapopulations defined by asymmetric homophily
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability
    roles: writing, model development
    corresponding: true
    email: maturner@stanford.edu
bibliography: [../library.bib]
# bibliography: this.bib
classoption: abstract
keep-tex: true
abstract: |
  \noindent hello
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
---

## Introduction

To equitably prepare for pandemics in our uncertain world we need to [@Rodo2021; @Adashi2022]. The global COVID-19 pandemic highlighted the current lack of outcome equality, where a country's inequality (measured by the Gini index) predicted higher burden of COVID-19 infections [@Su2022]. Developing computational models can help us understand basic behavioral and social factors underlying the relationship between behavior and disease, but need to be applied to real-world situations because, first and foremost, there are vulnerable populations who need effective mitigation plans for the inevitable next pandemic. Second, a generic, yet predictive and application-ready, framework for behavioral-epidemiological metapopulation models can only be developed through making contact with real-world scenarios to hone the model, identifying missing or unnecessary components, and building a repertoire or menu of potential modeling choices that may be important. By applying abstract mechanistic models to high-priority real-world problems, we will develop a more broadly useful science of pandemic preparedness.

In this work we contribute to this program of more equitable pandemic preparedness by developing a compartmental behavioral-epidemiological metapopulation model of COVID-19 transmission between a Brazilian city, Dourados, in Mato Grosso do Sul, and the neighboring Dourados Indigenous Reserve [@DeOliveira2023]. De Oliveira and co-authors hypothesized that COVID-19 variant diffusion from Dourados City into the Dourados Reserve is driven by young Indigenous men traveling between the reserve and the city of Dourados for economic reasons. COVID-19 burden was felt disproportionately by Indigenous peoples in Brazil [@Simionatto2020], but may have been worse in the Dourados Reserve due to its immediate proximity to the Dourados. There are many complicating factors that make it difficult to evaluate this hypothesis directly without adding some additional qualifications. Here we develop a mechanistic model of epidemiological transmission with asymmetric social mixing that evaluates how best to model group structure in the Reserve. In turn, this model is used to evaluate potential economic interventions that target aid to different groups to reduce infections in the reserve.

-   De Oliveira and colleagues tracked the spread of different SARS-CoV-2 variants by village and by variant.

![Time series of COVID-19 spread through the Dourados Indigenous Reserve, with significant events highlighted. The first B.1.1 variant cases were introduced by the Guarani ethnicity (1), followed about one month later by a rapid rise in cases in Jaguapiru among the Terena ethnicity(2); several months later a rapid rise in Bororó was caused by a rapid rise among Guarani (3). Zeta was introduced at the end of 2020 (4), with a rapid rise around the holiday season especially in Jaguapiru among Guarani (5); later in February cases rose in Bororó among the Terena (6).](Figures/Time%20Series.pdf){width=80%}

-   The goal of our model is two-fold. First, to identify parameters that could re-create the time series plots in Figure 1.
-   Second, we aim to use a parameterized model to understand how different intervention strategies could reduce infections among Reserve residents. For example, perhaps since Terena are suspected to travel to the 
-   While we know a lot about transmission in the reserve, and the work of de Oliveira and colleagues advanced our understanding of transmission in the Reserve, we do not have reliable estimates of exactly how many cases there were per ethnicity and village. We also have only word-of-mouth information about the frequency of travel to the city for each ethnicity and location. Our model motivates the need for a more thorough understanding of behavior and epidemiological parameters in the Dourados Indigenous Reserve, and points the way towards future research in other Indigenous communities, and more complex situations.

## Model

![We can model the relevant groups for mixing as either village residence or ethnicity. Transmission, mixing, and mobility parameters would be set accordingly. The $C_g$ parameters would be reduced if economic aid were allocated to keep group $g$ on the Reserve during outbreaks.](Figures/ModelSketch.pdf){fig-align="center"}

## Analysis

## Discussion